Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election 

### 2024-04-18 paper/paper(1).qmd
Add files via upload

```{r setup, include=FALSE, message=FALSE}
## Import necessary libraries
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(readr)
library(knitr)
library(kableExtra)
library(rstanarm)
library(arrow)
library(tidybayes)
library(bayesplot)
```

@@

## Hypothesis
To enhance control and prevention of illegal immigration, the Liberal Party has suggested modifications to the Immigration and Refugee Protection Act. They propose establishing an annual refugee quota of 250, specifically aimed at supporting human rights defenders, journalists, and humanitarian workers seeking safety and resettlement in Canada. Looking ahead, the Liberal Party intends to incrementally increase this refugee quota, aiming for refugees to comprise 15% of all immigrants. Conversely, the share of economic immigrants is planned to decrease to 58%, while the family reunification immigrant ratio will remain constant. These proportions are expected to be maintained over the next two years. Moreover, the Canadian national government recently unveiled an ambitious plan to admit 500,000 migrants annually by 2025, looking for nearly 1.5 million newcomers in Canada within the next three years (Canada's Immigration Levels Plan 2024-2026). Historically, the Liberal government's immigration strategy tends to be more permissive, offering broader entry routes compared to that of the Conservative Party[@Nakhaie_2006].
Based on our analysis of historical trends, governmental policies, and electoral procedures, our hypothesis aligns with the prevailing expectation that the Conservatives will emerge as the likely winners in the 2025 election.
**We anticipate that the Liberals will likely secure the highest number of votes, with the Conservatives expected to closely follow in second place.** Our research indicates that the vote-counting process in Canada significantly influences the outcome, as voters do not directly elect a specific prime minister candidate.
Initially, all votes are tallied based on first preferences.
If a candidate obtains 50% of the vote, they win.
If not, the candidate with the fewest votes is eliminated[@Elections_2023].
Then, votes from supporters of the eliminated candidate are redistributed to their second-choice candidate.
This iterative redistribution continues until one candidate achieves 50% of the vote.
Hence, our objective is to forecast the election probabilities for the Conservative and Liberal parties and compare which is higher[@Elections_2023].

@@

**Census Data**: Census data refer to information systematically gathered from the entire population of a defined area at a specific point in time.
The census data we have comes from the General Social Survey (GSS), conducted on August 12, 2022.
We've made necessary adjustments to ensure it matches our survey data.
Data collection aims to provide a detailed snapshot of demographic, economic, and social characteristics.The GSS is a nationwide survey program designed to collect data on societal trends, behaviors, and attributes across the entire population, offering insights into the changing dynamics of communities.

**Survey Data**: Survey data, are collected from a sample of the population rather than every individual.
This method is often used when it's impractical or unnecessary to include everyone.
Survey data aim to infer the characteristics of the larger population based on the responses of the sample.
The survey data, obtained from the Canada Election Study (CES) in 2023, includes responses from over 37,000 participants.
The CES focuses specifically on voter behavior, attitudes, and the electoral process within the Canadian context.
Unlike the GSS, which aims for a comprehensive overview, the CES targets specific topics of interest to derive insights from a segment of the population.

@@

### Data Summary Measures
```{r}
#| label: tbl-data-summary-numeric
#| tbl-cap: Summary of Numerical Variables Across Census and Survey Data
#| echo: false
numeric_summary_df <- data.frame(
  Variable = c("Age in Census Data", "Age in Survey Data", 
               "Liberal Votes in Survey", "Conservative Votes in Survey"),
  Minimum = c(15.0, 18.0, 0.000, 0.000),
  Q1 = c(37.0, 36.0, 0.000, 0.000),
  Median = c(54.0, 53.0, 0.000, 0.000),
  Average = c(52.18, 51.3, 0.267, 0.249),
  Q3 = c(67.0, 66.0, 1.000, 0.000),
  Maximum = c(80.0, 97.0, 1.000, 1.000)
)

@@

### Data Visualization

@@

@fig-age-distribution displays the age distribution from both census (in red) and survey (in blue) datasets.
The census distribution exhibits a unimodal peak around the 50-60 year age range and a sharp decline beyond 80 years, suggesting a concentration of middle to late-middle-aged respondents.
Conversely, the survey dataset portrays a wider distribution, extending towards older ages with a noticeable skew to the right, indicating the presence of older individuals up to 97 years old.
The regions of overlap appear purple, indicating age ranges common to both datasets, with the census data generally showing higher frequencies in most age categories.
This comparison highlights the distinct age profiles captured by the two sources, with a significant representation of middle-aged individuals in both, with the survey dataset capturing a broader demographic.
\newpage

@@

@fig-province-distribution contrasts the provincial distribution of respondents from both survey and census datasets, with provinces along the x-axis and respondent counts on the y-axis.
The survey data, shown in blue, reaches its highest point in Ontario, indicating a significantly larger number of respondents compared to other provinces in the survey.
This peak towers above all other survey data points and also surpasses the census data count for Ontario, which suggests a disproportionate representation or potentially an over-sampling in the survey for Ontario.
In contrast, the lowest point for the survey dataset appears in Prince Edward Island, with very few respondents, paralleled by a similarly low count for this province in the census dataset, shown in red.
Notably, the census data peaks in Quebec, where the red bar indicates the highest frequency among all provinces for the census, while Alberta's census count is at its lowest, barely rising above the baseline.
This visual comparison reveals not only the variations in population coverage between the two types of data but also significant regional differences within the datasets, which could reflect varied sampling methods or demographic distributions.
\newpage

@@

This discrepancy highlights potential differences in sampling methods, response rates, or the representativeness of each dataset relative to the underlying population.
The most pronounced difference is at the "Limited Education" level, where the Survey dataset shows significantly lower counts compared to the Census dataset, indicating a possible overrepresentation of individuals with limited education in the censaus or an underrepresentation in the survey.
The data suggest that demographic, socioeconomic, or regional factors may influence the education level distribution in each dataset, and further statistical analysis could be warranted to determine the significance and cause of these differences.
The differences observed in the chart might also reflect methodological variations in data collection or inherent biases within the sampling frames used.
\newpage

@@

@fig-religion-distribution delineates the reported religious affiliations in both the census and survey datasets, differentiated by color intensity, with the census depicted in a darker red and the survey in a lighter pink.
The x-axis is categorized into 'No' for non-affiliated individuals and 'Yes' for those with religious affiliation.
A striking contrast is observed in the 'Yes' category, where the census data’s count towers over that of the survey, indicating a greater number of religiously affiliated respondents.
Conversely, for those reporting no religious affiliation, the survey's lighter pink bar is comparable in height to the census's, suggesting a more even distribution between the two datasets.
This discrepancy underscores a pronounced difference in the representation of religious affiliation between the two data sources, potentially alluding to divergent demographic characteristics or response patterns within the datasets.
\newpage

@@

@fig-voting-preferences contrasts the voting preferences for the Conservative and Liberal parties, delineating counts of participants who voted (blue bars) against those who did not (red bars).
Notably, there is a substantial number of non-voters in the Conservative category, exceeding their Liberal counterparts, suggesting a larger disenchantment or mobilization challenge within the Conservative base.
Despite this, the voter turnout (blue bars) is fairly balanced between the parties, indicating a competitive engagement among those who participated in the vote.
The overall higher presence of red bars signifies a general trend of voter disengagement, which may warrant further investigation through regression analysis and post-stratification to grasp the full scope of electoral dynamics and the broader implications for the Canadian democratic process.
\newpage

@@

# Model
We utilized the rstanarm package for constructing the generalized linear model[@R-rstanarm], the rstanarm package serves as a user-friendly interface in R for fitting Bayesian regression models using the Stan probabilistic programming language.
It enables users to effortlessly construct various types of regression models, including linear, logistic, and Poisson regression, while leveraging the power of Bayesian methods for inference.
By seamlessly integrating with Stan, rstanarm provides efficient algorithms for model fitting and accurate estimates of uncertainty.
Additionally, the package offers diagnostic tools for assessing model convergence and sensitivity to prior specifications, making it an invaluable resource for Bayesian data analysis in R.
\newpage

@@

-   $y_i$: Binary outcome of voting for the Liberal/Conservative party (1) or not (0)
-   $\pi_i$: Probability of voting for the Liberal/Conservative party
```{=tex}
\begin{align}
y_i | \pi_i &\sim \text{Bernoulli}(\pi_i) \\
\text{logit}(\pi_i) &= \beta_0 + \alpha^{age}_{a[i]} + \alpha^{gender}_{g[i]} + \alpha^{province}_{p[i]} + \alpha^{religion}_{r[i]} + \alpha^{income}_{m[i]} + \alpha^{education}_{e[i]} \\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\alpha_{a}^{age} &\sim \text{Normal}(0, \sigma^{2}_{age}) \quad \text{for } a \text{ in } 1, 2, \ldots, A \\
\alpha_{g}^{gender} &\sim \text{Normal}(0, \sigma^{2}_{gender}) \quad \text{for } g \text{ in } Male,Female \\
\alpha_{p}^{province} &\sim \text{Normal}(0, \sigma^{2}_{province}) \quad \text{for } p \text{ in } BC, AB, MB, NB, NL, NS, ON, PE, QC, SK \\
\alpha_{r}^{religion} &\sim \text{Normal}(0, \sigma^{2}_{religion}) \quad \text{for } r \text{ in } YES,NO \\
\alpha_{m}^{income} &\sim \text{Normal}(0, \sigma^{2}_{income}) \quad \text{for } m \text{ in } Lower Class, Middle Class, Upper Class\\
\alpha_{e}^{education} &\sim \text{Normal}(0, \sigma^{2}_{education}) \quad \text{for } e \text{ in } Low Education, Some Education, High Education \\
\sigma^{2}_{age} &\sim \text{Exponential}(1) \\
\sigma^{2}_{gender} &\sim \text{Exponential}(1) \\
\sigma^{2}_{province} &\sim \text{Exponential}(1) \\
\sigma^{2}_{religion} &\sim \text{Exponential}(1) \\
\sigma^{2}_{income} &\sim \text{Exponential}(1) \\
\sigma^{2}_{education} &\sim \text{Exponential}(1) 
\end{align}
```

## Model justification

## Model Coefficients Interpretation

@@

$$log(\frac{p}{1-p}) = \beta_0+\beta_1x_{age}+\beta_2x_{gender}+\beta_3x_{porvince}+\beta_{4}x_{religion}+\beta_{5}x_{income}+\beta_{6}x_{education}$$

$p$ is the probability of an individual casting their vote for a certain political party (Liberal or Conservative).

@@

install.packages("MASS")
# Load the necessary library
library(MASS)
library(broom)
library(knitr)

# Fit the initial linear model
initial_model <- lm(vote_conservative ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)

# Perform stepwise model selection using stepAIC
step_cons <- stepAIC(initial_model, direction="both", trace = FALSE)

# Use broom::tidy to get a tidy data frame of the model summary with confidence intervals
tidy_cons <- broom::tidy(step_cons, conf.int = TRUE)

# Select the columns you want for the table, excluding statistic and p-value
tidy_cons_selected <- tidy_cons[, c("term", "estimate", "std.error", "conf.low", "conf.high")]

# Create the table with knitr::kable()
kable_output <- knitr::kable(tidy_cons_selected, format = "html", caption = "Model Summary for Conservative Vote Prediction with Confidence Intervals")

# Display the table output
kable_output

@@

@@

# Assuming 'vote_liberal' is the column in your 'cleaned_survey_data' that represents voting for liberal
initial_model_lib <- lm(vote_liberal ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)
step_lib <- stepAIC(initial_model_lib, direction="both", trace = FALSE)

@@

kable_output_lib <- knitr::kable(tidy_lib_selected, format = "html", caption = "Model Summary for Liberal Vote Prediction with Confidence Intervals")
kable_output_lib

@@

@tbl-coef-summary-liberal

`

@tbl-coef-summary-conservative

@tbl-ypshat-summary
```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
if (!require(MASS)) {
    install.packages("MASS")
    library(MASS)
} else {
    library(MASS)
}

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
Party_Name <- c("Conservative", "Liberal")
yPS_primary <- c(cons_estimate,lib_estimate)
yPS_AIC <- c(cons_aic ,lib_aic)

knitr::kable(tibble(Party_Name, yPS_primary, yPS_AIC), caption = "The Hat of yPS For Each Party, Between Primary Model and Final Model")

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)
library(broom)

required_vars <- c("vote_liberal", "age", "sex", "province", "education", "religion", "income_before_tax")
missing_vars <- setdiff(required_vars, names(cleaned_survey_data))

@@

ggplot(model_summary_lib, aes(x = estimate, y = reorder(term, estimate))) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Estimate", y = "Coefficient", title = "Coefficient Estimates for Liberal Vote Prediction") +
  theme_minimal()

@@

@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)

# Assuming 'cleaned_survey_data' contains 'province' and 'vote_liberal',
# where 'vote_liberal' is a binary (0/1) indicating whether a vote for liberal was cast.

# Calculate proportions and confidence intervals by province for the liberal party
liberal_votes_by_province <- cleaned_survey_data %>%
  group_by(province) %>%
  summarise(
    proportion = mean(vote_liberal, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    .groups = 'drop'  # Drop grouping for further manipulation if needed
  ) %>%
  ungroup()  # Ensure we remove the grouping for the ggplot

# Create the plot for the liberal party
ggplot(liberal_votes_by_province, aes(x = reorder(province, -proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  # Flip the axes to have provinces on the y-axis
  labs(x = "Proportion", y = "Province", title = "Estimated Proportion of Voting for Liberal by Province") +
  theme_minimal()

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)

# Assuming 'cleaned_survey_data' is your dataframe and it contains the variables 'education' and 'vote_liberal'
# where 'vote_liberal' is a binary (0/1) indicating whether a vote for liberal was cast.

liberal_votes_by_education <- cleaned_survey_data %>%
  group_by(education) %>%
  summarise(
    proportion = mean(vote_liberal, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    .groups = 'drop'  # This will automatically ungroup the data (requires dplyr version 1.0.0 or higher)
  ) %>%
  arrange(desc(proportion))

ggplot(liberal_votes_by_education, aes(x = reorder(education, proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  # Flip the axes to have education levels on the y-axis
  labs(x = "Proportion", y = "Education", title = "Proportion of Votes for Liberal by Education Level") +
  theme_minimal()

## Plot of binary logistic regression models

# Discussion
Recent research highlights that the primary concerns for Canadian voters leading up to the 2025 election include the rising cost of living, housing, healthcare, the economy, and the environment. These issues are pivotal to the electorate, and the political parties’ stances on them are under intense scrutiny. Our analysis, drawn from five different polls, suggests a marginal preference for the Conservative Party over the Liberal Party. This trend is echoed in Barik's 2023 article, which indicates a public perception of the Conservative Party's superior handling of these key issues. Thus, our working hypothesis posits that the Conservative Party holds a slender advantage in the race for the upcoming election[@Barik_2023].

>>>>

```{r setup, include=FALSE, message=FALSE}
## Import necessary libraries
install.packages("MASS")
library(ggplot2)
library(dplyr)
library(tidyverse)
library(here)
library(readr)
library(knitr)
library(kableExtra)
library(rstanarm)
library(arrow)
library(tidybayes)
library(bayesplot)
library(MASS)
library(broom)
```

@@

## Hypothesis

To enhance control and prevention of illegal immigration, the Liberal Party has suggested modifications to the Immigration and Refugee Protection Act.
They propose establishing an annual refugee quota of 250, specifically aimed at supporting human rights defenders, journalists, and humanitarian workers seeking safety and resettlement in Canada.
Looking ahead, the Liberal Party intends to incrementally increase this refugee quota, aiming for refugees to comprise 15% of all immigrants.
Conversely, the share of economic immigrants is planned to decrease to 58%, while the family reunification immigrant ratio will remain constant.
These proportions are expected to be maintained over the next two years.
Moreover, the Canadian national government recently unveiled an ambitious plan to admit 500,000 migrants annually by 2025, looking for nearly 1.5 million newcomers in Canada within the next three years (Canada's Immigration Levels Plan 2024-2026).
Historically, the Liberal government's immigration strategy tends to be more permissive, offering broader entry routes compared to that of the Conservative Party[@Nakhaie_2006].
Based on our analysis of historical trends, governmental policies, and electoral procedures, our hypothesis aligns with the prevailing expectation that the Conservatives will emerge as the likely winners in the 2025 election.
**We anticipate that the Liberals will likely secure the highest number of votes, with the Conservatives expected to closely follow in second place.** Our research indicates that the vote-counting process in Canada significantly influences the outcome, as voters do not directly elect a specific prime minister candidate.
Initially, all votes are tallied based on first preferences.
If a candidate obtains 50% of the vote, they win.
If not, the candidate with the fewest votes is eliminated[@Elections_2023].
Then, votes from supporters of the eliminated candidate are redistributed to their second-choice candidate.
This iterative redistribution continues until one candidate achieves 50% of the vote.
Hence, our objective is to forecast the election probabilities for the Conservative and Liberal parties and compare which is higher[@Elections_2023].

@@

**Census Data**: Census data refer to information systematically gathered from the entire population of a defined area at a specific point in time.
The census data we have comes from the General Social Survey (GSS), conducted on August 12, 2022.
We've made necessary adjustments to ensure it matches our survey data.
Data collection aims to provide a detailed snapshot of demographic, economic, and social characteristics.The GSS is a nationwide survey program designed to collect data on societal trends, behaviors, and attributes across the entire population, offering insights into the changing dynamics of communities[@GSS_2022].

**Survey Data**: Survey data, are collected from a sample of the population rather than every individual.
This method is often used when it's impractical or unnecessary to include everyone.
Survey data aim to infer the characteristics of the larger population based on the responses of the sample.
The survey data, obtained from the Canada Election Study (CES) in 2023, includes responses from over 37,000 participants.
The CES focuses specifically on voter behavior, attitudes, and the electoral process within the Canadian context.
Unlike the GSS, which aims for a comprehensive overview, the CES targets specific topics of interest to derive insights from a segment of the population[@CES_2023].

@@

### Data Summary Measures

```{r}
#| label: tbl-data-summary-numeric
#| tbl-cap: Summary of Numerical Variables Across Census and Survey Data
#| echo: false
numeric_summary_df <- data.frame(
  Variable = c("Age in Census Data", "Age in Survey Data", 
               "Liberal Votes in Survey", "Conservative Votes in Survey"),
  Minimum = c(15.0, 18.0, 0.000, 0.000),
  Q1 = c(37.0, 36.0, 0.000, 0.000),
  Median = c(54.0, 53.0, 0.000, 0.000),
  Average = c(52.18, 51.3, 0.267, 0.249),
  Q3 = c(67.0, 66.0, 1.000, 0.000),
  Maximum = c(80.0, 97.0, 1.000, 1.000)
)

@@

\newpage
### Data Visualization

@@

@fig-age-distribution displays the age distribution from both census (in red) and survey (in blue) datasets.
The census distribution exhibits a unimodal peak around the 50-60 year age range and a sharp decline beyond 80 years, suggesting a concentration of middle to late-middle-aged respondents.
Conversely, the survey dataset portrays a wider distribution, extending towards older ages with a noticeable skew to the right, indicating the presence of older individuals up to 97 years old.
The regions of overlap appear purple, indicating age ranges common to both datasets, with the census data generally showing higher frequencies in most age categories.
This comparison highlights the distinct age profiles captured by the two sources, with a significant representation of middle-aged individuals in both, with the survey dataset capturing a broader demographic.

@@

@fig-province-distribution contrasts the provincial distribution of respondents from both survey and census datasets, with provinces along the x-axis and respondent counts on the y-axis.
The survey data, shown in blue, reaches its highest point in Ontario, indicating a significantly larger number of respondents compared to other provinces in the survey.
This peak towers above all other survey data points and also surpasses the census data count for Ontario, which suggests a disproportionate representation or potentially an over-sampling in the survey for Ontario.
In contrast, the lowest point for the survey dataset appears in Prince Edward Island, with very few respondents, paralleled by a similarly low count for this province in the census dataset, shown in red.
Notably, the census data peaks in Quebec, where the red bar indicates the highest frequency among all provinces for the census, while Alberta's census count is at its lowest, barely rising above the baseline.
This visual comparison reveals not only the variations in population coverage between the two types of data but also significant regional differences within the datasets, which could reflect varied sampling methods or demographic distributions.

@@

This discrepancy highlights potential differences in sampling methods, response rates, or the representativeness of each dataset relative to the underlying population.
The most pronounced difference is at the "Limited Education" level, where the Survey dataset shows significantly lower counts compared to the Census dataset, indicating a possible overrepresentation of individuals with limited education in the censaus or an underrepresentation in the survey.
The data suggest that demographic, socioeconomic, or regional factors may influence the education level distribution in each dataset, and further statistical analysis could be warranted to determine the significance and cause of these differences.
The differences observed in the chart might also reflect methodological variations in data collection or inherent biases within the sampling frames used.

@@

@fig-religion-distribution delineates the reported religious affiliations in both the census and survey datasets, differentiated by color intensity, with the census depicted in a darker red and the survey in a lighter pink.
The x-axis is categorized into 'No' for non-affiliated individuals and 'Yes' for those with religious affiliation.
A striking contrast is observed in the 'Yes' category, where the census data’s count towers over that of the survey, indicating a greater number of religiously affiliated respondents.
Conversely, for those reporting no religious affiliation, the survey's lighter pink bar is comparable in height to the census's, suggesting a more even distribution between the two datasets.
This discrepancy underscores a pronounced difference in the representation of religious affiliation between the two data sources, potentially alluding to divergent demographic characteristics or response patterns within the datasets.

@@

@fig-voting-preferences contrasts the voting preferences for the Conservative and Liberal parties, delineating counts of participants who voted (blue bars) against those who did not (red bars).
Notably, there is a substantial number of non-voters in the Conservative category, exceeding their Liberal counterparts, suggesting a larger disenchantment or mobilization challenge within the Conservative base.
Despite this, the voter turnout (blue bars) is fairly balanced between the parties, indicating a competitive engagement among those who participated in the vote.
The overall higher presence of red bars signifies a general trend of voter disengagement, which may warrant further investigation through regression analysis and post-stratification to grasp the full scope of electoral dynamics and the broader implications for the Canadian democratic process.

@@

# Model

We utilized the rstanarm package for constructing the generalized linear model[@R-rstanarm], the rstanarm package serves as a user-friendly interface in R for fitting Bayesian regression models using the Stan probabilistic programming language.
It enables users to effortlessly construct various types of regression models, including linear, logistic, and Poisson regression, while leveraging the power of Bayesian methods for inference.
By seamlessly integrating with Stan, rstanarm provides efficient algorithms for model fitting and accurate estimates of uncertainty.
Additionally, the package offers diagnostic tools for assessing model convergence and sensitivity to prior specifications, making it an invaluable resource for Bayesian data analysis in R.
\newpage

@@

-   $y_i$: Binary outcome of voting for the Liberal/Conservative party (1) or not (0)
-   $\log(\frac{\pi_i}{1-\pi_i})$: Probability of voting for the Liberal/Conservative party

```{=tex}
\begin{align}
y_i | \pi_i &\sim \text{Bernoulli}(\pi_i) \\
\text{logit}(\frac{\pi_i}{1-\pi_i}) &= \beta_0 + \beta_1 x_{i,\text{age}} + \beta_2 x_{i,\text{gender}} + \beta_3 x_{i,\text{province}} + \beta_4 x_{i,\text{religion}} + \beta_5 x_{i,\text{income}} + \beta_6 x_{i,\text{education}} \\
\beta_0 &\sim \text{Normal}(0, 1) \\
\beta_1, \beta_2, \beta_3, \beta_4, \beta_5, \beta_6 &\sim \text{Normal}(0, 1)
\end{align}
```
The model, as formulated in the Bayesian logistic regression framework, is designed to examine the impact of various sociodemographic factors on the probability of an individual voting for a specific political party, such as the Liberals or Conservatives in Canada.
The probability $\log(\frac{\pi_i}{1-\pi_i})$of a particular voting outcome follows a Bernoulli distribution, where each individual's probability of voting for the party is modeled using logistic regression.

## Components of the Model

**Bernoulli Distribution** Each binary outcome $y_i$ (whether an individual votes for the party or not) follows a Bernoulli distribution with success probability $\log(\frac{\pi_i}{1-\pi_i})$, which is the probability that $y_i = 1$.

**Logit Link Function** The logistic function or logit link transforms the linear combination of predictors into a probability (0 to 1).
It is defined as $\log(\frac{\pi_i}{1-\pi_i})$, mapping real numbers from the linear predictors to the interval $[0,1]$.

**Predictors and Coefficients** - $\beta_0$ (Intercept): Represents the baseline log-odds of voting for the party when all predictor variables are at zero or their reference categories.
- $\beta_1$ (Age): Reflects the influence of age on the probability of voting for the party.
A positive $\beta_1$ indicates increasing age is associated with a higher likelihood of voting for the party.
- $\beta_2$(Gender): Adjusts the probability based on gender, typically comparing males to females if females are the reference category.
- $\beta_3$(Province): Accounts for regional variations in voting patterns, with coefficients for each province relative to a reference province.
- $\beta_4$ (Religion): Represents the difference in voting probability between individuals identifying with a religion versus those who do not.
- $\beta_5$ (Income): Captures the effect of income on voting behavior.
- $\beta_6$ (Education): Reflects how varying levels of education influence voting decisions.

**Prior Distributions** All coefficients, including the intercept, are assumed to follow a Normal(0, 1) distribution as priors.
These priors are weakly informative, centering the expected value of each effect around zero but allowing the data to inform the final posterior distributions.
This approach helps in regularizing the estimates, particularly in complex models or when data are limited.

The model effectively quantifies how demographic and socioeconomic factors influence political behavior using Bayesian inference.
This method provides a probabilistic framework that not only gives estimates of the effects but also accounts for uncertainty in a coherent way, offering more flexibility and insight than traditional frequentist approaches.
By assuming normal priors, it indicating a prior belief that none of the predictors has an unusually large effect, but remain open to being informed by the data collected.

@@

$$log(\frac{p}{1-p}) = \beta_0+\beta_1x_{age}+\beta_2x_{gender}+\beta_3x_{porvince}+\beta_{4}x_{religion}+\beta_{5}x_{income}+\beta_{6}x_{education}$$ **Model Coefficients Interpretation** $p$ is the probability of an individual casting their vote for a certain political party (Liberal or Conservative).

@@

library(MASS)
library(broom)
library(knitr)

initial_model <- lm(vote_conservative ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)

step_cons <- stepAIC(initial_model, direction="both", trace = FALSE)  
tidy_cons <- broom::tidy(step_cons, conf.int = TRUE)

tidy_cons_selected <- tidy_cons[, c("term", "estimate", "std.error", "conf.low", "conf.high")]

kable_output <- knitr::kable(tidy_cons_selected, format = "html")

kable_output

@@

@tbl-coef-summary-conservative presents calculated coefficients that integrate into our logistic regression model for conservative party, derived from the survey dataset from General Social Survey[@GSS_2022]. The `knitr` package [@rKnitr] was utilized to generate the table. These coefficients represent the adjusted effects of various predictors such as age, province, education, religion, and income brackets on the log odds of an individual reporting a certain behavior or characteristic.

In the context of R Studio, the `MASS` package[@MASS], although not directly used here, is often employed for its functions related to linear and quadratic discriminant function analysis, while the `broom` package tidies the output of statistical tests into data frames to facilitate further data analysis or visualization. Specifically, `broom` [@broom]can convert the summary output of regression models into an easy-to-handle data frame, as shown in the table.

The table details the point estimates for each predictor along with their standard errors and confidence intervals, providing insight into the precision of the estimates and the potential range within which the true value of the coefficients may lie. Notably, the intercept and coefficients for variables such as 'provincePrince Edward Island' and 'income_before_taxUpper Class' have both negative and positive values, indicating their varied influence on the outcome variable. The confidence intervals suggest the level of certainty about the estimates, where a narrower interval represents a more precise estimate. These results contribute to our understanding of the factors influencing voter behavior, guiding the development of targeted strategies for engagement and policy-making.

@@

initial_model_lib <- lm(vote_liberal ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)
step_lib <- stepAIC(initial_model_lib, direction="both", trace = FALSE)

@@

kable_output_lib <- knitr::kable(tidy_lib_selected, format = "html")
kable_output_lib

@@

@tbl-coef-summary-liberal provided outlines the logistic regression model coefficients pertinent to predicting voting behavior for the Liberal Party, based on the survey data extracted from the General Social Survey [@GSS_2022]. Fabricated through the adept use of the `knitr` package [@rKnitr], the table concisely encapsulates the nuanced influence of various demographic and socioeconomic predictors — age, gender, province, educational, religion, and income — on the likelihood of an individual voting Liberal.

Diving into the details, the table lists the estimated coefficients and their accompanying standard errors, offering a statistical measure of the precision of the estimates. The confidence intervals provided for each coefficient — lower (conf.low) and higher (conf.high) bounds — delineate the range in which the true coefficient values are likely to fall, with tighter intervals indicating higher precision. For instance, the negative coefficient for 'provincePrince Edward Island' suggests a lesser inclination to vote Liberal in that region compared to the reference province, while the positive coefficients for 'provinceNewfoundland and Labrador' and 'provinceOntario' reflect a stronger propensity for Liberal support.

While not employed directly in the analysis leading to this table, the `MASS` package [@MASS] is typically revered for its robust functionalities in linear discriminant analysis, an alternative approach to examining group differences. Complementing this, the `broom` package [@broom] elegantly streamlines the model outputs into tidy data frames, greatly simplifying the transition from statistical testing to interpretable results and potential graphical representation.

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
if (!require(MASS)) {
    install.packages("MASS")
    library(MASS)
} else {
    library(MASS)
}

@@

```{r, label="tbl-yps-summary", echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-yps-summary
#| tbl-cap: The Hat of yPS For Each Party, Between Primary Model and Final Model
#| echo: false
Party_Name <- c("Conservative", "Liberal")
yPS_primary <- c(cons_estimate,lib_estimate)
yPS_AIC <- c(cons_aic ,lib_aic)

knitr::kable(tibble(Party_Name, yPS_primary, yPS_AIC))

@@

Based on @tbl-yps-summary, the estimates suggest differing probabilities for each major party.
For the Conservative Party, the final model adjustment indicates an estimated voting probability of about 28.66%.
This projection implies that roughly 28.66% of the population, characterized by certain age, province, education, religion, and income factors, would likely vote Conservative in the 2025 Canadian Election.
In contrast, the Liberal Party's estimated voting probability is slightly higher at approximately 28.83% under the AIC-adjusted model, suggesting that nearly 28.83% of individuals with the specified demographics are predicted to cast their vote for the Liberals.
\newpage

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#| label: fig-liberal-coefficients
#| tbl-cap: "Coefficient Estimates for Liberal Vote Prediction"
#| echo: false
required_vars <- c("vote_liberal", "age", "sex", "province", "education", "religion", "income_before_tax")
missing_vars <- setdiff(required_vars, names(cleaned_survey_data))

@@

ggplot(model_summary_lib, aes(x = estimate, y = reorder(term, estimate))) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Estimate", y = "Coefficient") +
  theme_minimal()

@@

@fig-liberal-coefficients displays the logistic regression model's estimated coefficients relating to the likelihood of supporting the Liberal party, based on various factors. Each point (dot) on the plot signifies the model's coefficient for a specific predictor, with the position on the x-axis indicating the magnitude and direction of the predictor's effect. Horizontal lines represent the confidence intervals, giving a visual representation of the estimate's precision and the range within which we can be confident that the actual value of the coefficient lies.

The intercept, situated near zero, represents the baseline log-odds of voting for the Liberal party when all predictors are at their reference levels. The estimates for provinces such as Newfoundland and Labrador, Ontario, and New Brunswick are situated to the right of the intercept, demonstrating a positive relationship with Liberal support. Conversely, the coefficient for Prince Edward Island is markedly left of the intercept, suggesting a negative association with the likelihood of voting for the Liberals.

Notably, the coefficient for 'sexMale' is negative, indicating that being male is associated with a lower likelihood of voting for the Liberal party compared to the reference group, presumably females. Age shows a small positive coefficient, hinting at a slight increase in Liberal support with advancing age.

'Upper Class' and 'Middle Class' both have positive coefficients, suggesting higher income brackets may correlate with increased support for the Liberal party. In terms of education, 'Low Education' and 'Some Education' have negative coefficients, indicating these groups might be less inclined to vote Liberal.

The red dashed line at zero is a reference that indicates no effect; coefficients to the right demonstrate a positive influence on Liberal support, while those to the left indicate a negative influence. The plot orders the variables on the y-axis, providing a hierarchy of impact, from the most positive at the top to the most negative at the bottom.

This figure offers a nuanced understanding of how different demographic and socioeconomic variables might influence Liberal party support, informing strategic political engagement and policy development. The spread of the confidence intervals highlights areas of greater or lesser certainty in the model's predictions, essential for interpreting the robustness of the variables' effects.

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#| label: fig-liberal-votes-by-province
#| tbl-cap: "Estimated Proportion of Voting for Liberal by Province"
#| echo: false
liberal_votes_by_province <- cleaned_survey_data %>%
  group_by(province) %>%
  summarise(
    proportion = mean(vote_liberal, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    .groups = 'drop'  
  ) %>%
  ungroup()

ggplot(liberal_votes_by_province, aes(x = reorder(province, -proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  
  labs(x = "Proportion", y = "Province") +
  theme_minimal()

```
@fig-liberal-votes-by-province visualizes the estimated proportions of support for the Liberal party across various Canadian provinces. Each dot on the graph indicates the mean proportion of Liberal votes within a province, while the horizontal lines represent the error bars, which detail the confidence interval for these proportions. The span of each line reflects the level of uncertainty or variability in the estimate—wider lines suggest more uncertainty.
